Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-15 paper.tex
Switch to minted.

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{xcolor}
\usepackage{xspace}

@@

As an example, let us consider a supervised learning task using logistic regression.
Given the API defined above, solving this problem is as simple as:
\begin{verbatim}
    from sklearn.linear_model import LogisticRegression

    clf = LogisticRegression(penalty="l1")
    clf.fit(X_train, y_train)
\end{verbatim}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"}, hence enabling
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C}, which
defines the strength of the regularization) are not explicitly defined and are
thus set to the default values. Upon calling \texttt{fit}, the model is then
learned from the training arrays \texttt{X\_train} and \texttt{y\_train}. Since
all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor. For example, to build a Random Forest on
the same training data, one would simply have to replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by \\
\texttt{RandomForestClassifier()}.

@@

The \textit{predictor} interface defines a \texttt{predict}
method taking as input some new data \texttt{X\_test} and producing as output
predictions for \texttt{X\_test}, based on the learned parameters of the
estimator (we call the input to \texttt{predict} ``\texttt{X\_test}'' in order
to emphasize that \texttt{predict} generalizes to new data). In the case of
supervised learning estimators, this method typically returns the predicted
labels or values computed by the model.  Continuing with the previous example,
predicted labels for \texttt{X\_test} can be obtained using the following
snippet:
\begin{verbatim}
    y_pred = clf.predict(X_test)
\end{verbatim}
Some unsupervised learning
estimators may however also implement that interface, as illustrated below:
\begin{verbatim}
    from sklearn.cluster import KMeans

    km = KMeans(n_clusters=10)
    km.fit(X_train)
    clust_pred = km.predict(X_test)
\end{verbatim}
In the above example, a \texttt{KMeans} estimator is first fit on the training
data \texttt{X\_train}. Then the \texttt{predict} method is called on the test
data \texttt{X\_test} in order to retrieve predicted cluster labels.
The \texttt{predict} method in \sklearn thus offers a unified way
to compute predictions with respect to unseen data.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection and dimensionality reduction
algorithms are all provided as transformers within the library.  In our example,
if one wanted to preprocess the inputs \texttt{X\_train} before fitting the
logistic regression estimator, e.g., by standardizing the feature values, one
could thus simply write:
\begin{verbatim}
    from sklearn.preprocessing import StandardScaler

    scaler = StandardScaler()
    scaler.fit(X_train)
    X_train = scaler.transform(X_train)
\end{verbatim}
Of course, in practice, it is important to apply the same preprocessing to the
test data \texttt{X\_test}. This can easily be done as follows:
\begin{verbatim}
    X_test = scaler.transform(X_test)
\end{verbatim}
Transformers also include a variety of learning algorithms, such as
dimension reduction (PCA, manifold learning techniques), kernel
approximation techniques, and any other approach that can be used to
derive a modified representation of the data.

Finally, let us also note that the \texttt{fit} method always returns the estimator
it was called on, hence making method chaining possible.  Using this technique,
we can thus rewrite the \texttt{StandardScaler} example above in a single line:
\begin{verbatim}
    X_train = StandardScaler().fit(X_train).transform(X_train)
\end{verbatim}
In addition, \texttt{fit(X\_train).transform(X\_train)} can further
be rewritten as \texttt{fit\_transform(X\_train)}.  Combining the \texttt{fit}
and \texttt{transform} methods into a single method \texttt{fit\_transform}
allows to share computations, which often makes a huge difference in practice.
In the same spirit, \sklearn sometimes provides the
\texttt{fit\_predict} method for combining the \texttt{fit} and \texttt{predict}
methods. This is especially useful for retrieving cluster predictions with
respect to the training data \texttt{X\_train} in clustering algorithms.

@@

As an example, the following snippet illustrates how to create a complex worflow
computing both linear PCA and kernel PCA components on \texttt{X\_train}
(through a \texttt{FeatureUnion}), selecting the 100 best of those features and
then feeding them to an $\ell_2$-regularized logistic regression model:
\begin{verbatim}
    from sklearn.pipeline import Pipeline
    from sklearn.decomposition import PCA, KernelPCA
    from sklearn.feature_selection import SelectKBest

    feat_union = FeatureUnion([('pca', PCA()),
                               ('kpca', KernelPCA(kernel="rbf"))])

    Pipeline([("pca", feat_union,
              ("feature_selection", SelectKBest(k=100)),
              ("logistic_model", LogisticRegression(penalty="l2"))
    ]).fit(X_train, y_train).predict(X_test)
\end{verbatim}

@@

As an example, the snippet below illustrates how to optimize the
hyper-parameters of an SVM classifier. It maximizes the $F_1$ score of the
classifier through a 10-fold cross-validation scheme.
\begin{verbatim}
    from sklearn.svm import SVC
    from sklearn.grid_search import GridSearchCV

    param_grid = [
      {'C': [1, 10, 100, 1000], 'kernel': ['linear']},
      {'C': [1, 10, 100, 1000], 'gamma': [0.001, 0.0001], 'kernel': ['rbf']},
    ]

    clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
    clf.fit(X_train, y_train)
    y_pred = clf.predict(X_test)
\end{verbatim}
Note that, in the above example, two distinct hyper-parameter grids are
considered. This is because it only makes sense to search for \texttt{gamma}
when \texttt{kernel} is set to \texttt{"rbf"}.  Additionally, we see that
\texttt{GridSearchCV} has a \texttt{predict} method, like any other classifier.
This is because \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} in fact
delegate the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
\texttt{score} methods to the best estimator, thus mimicking its behavior.

>>>>

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{minted}

\newminted{python}{frame=single}

% tt font with bold support
%\renewcommand{\ttdefault}{pcr}

@@

As an example, let us consider a supervised learning task using logistic regression.
Given the API defined above, solving this problem is as simple as:
\begin{pythoncode}
    from sklearn.linear_model import LogisticRegression

    clf = LogisticRegression(penalty="l1")
    clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"}, hence enabling
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C}, which
defines the strength of the regularization) are not explicitly defined and are
thus set to the default values. Upon calling \texttt{fit}, the model is then
learned from the training arrays \texttt{X\_train} and \texttt{y\_train}. Since
all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor. For example, to build a Random Forest on
the same training data, one would simply have to replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by \\
\texttt{RandomForestClassifier()}.

@@

The \textit{predictor} interface defines a \texttt{predict}
method taking as input some new data \texttt{X\_test} and producing as output
predictions for \texttt{X\_test}, based on the learned parameters of the
estimator (we call the input to \texttt{predict} ``\texttt{X\_test}'' in order
to emphasize that \texttt{predict} generalizes to new data). In the case of
supervised learning estimators, this method typically returns the predicted
labels or values computed by the model.  Continuing with the previous example,
predicted labels for \texttt{X\_test} can be obtained using the following
snippet:
\begin{pythoncode}
    y_pred = clf.predict(X_test)
\end{pythoncode}
Some unsupervised learning
estimators may however also implement that interface, as illustrated below:
\begin{pythoncode}
    from sklearn.cluster import KMeans

    km = KMeans(n_clusters=10)
    km.fit(X_train)
    clust_pred = km.predict(X_test)
\end{pythoncode}
In the above example, a \texttt{KMeans} estimator is first fit on the training
data \texttt{X\_train}. Then the \texttt{predict} method is called on the test
data \texttt{X\_test} in order to retrieve predicted cluster labels.
The \texttt{predict} method in \sklearn thus offers a unified way
to compute predictions with respect to unseen data.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection and dimensionality reduction
algorithms are all provided as transformers within the library.  In our example,
if one wanted to preprocess the inputs \texttt{X\_train} before fitting the
logistic regression estimator, e.g., by standardizing the feature values, one
could thus simply write:
\begin{pythoncode}
    from sklearn.preprocessing import StandardScaler

    scaler = StandardScaler()
    scaler.fit(X_train)
    X_train = scaler.transform(X_train)
\end{pythoncode}
Of course, in practice, it is important to apply the same preprocessing to the
test data \texttt{X\_test}. This can easily be done as follows:
\begin{pythoncode}
    X_test = scaler.transform(X_test)
\end{pythoncode}
Transformers also include a variety of learning algorithms, such as
dimension reduction (PCA, manifold learning techniques), kernel
approximation techniques, and any other approach that can be used to
derive a modified representation of the data.

Finally, let us also note that the \texttt{fit} method always returns the estimator
it was called on, hence making method chaining possible.  Using this technique,
we can thus rewrite the \texttt{StandardScaler} example above in a single line:
\begin{pythoncode}
    X_train = StandardScaler().fit(X_train).transform(X_train)
\end{pythoncode}
In addition, \texttt{fit(X\_train).transform(X\_train)} can further
be rewritten as \texttt{fit\_transform(X\_train)}.  Combining the \texttt{fit}
and \texttt{transform} methods into a single method \texttt{fit\_transform}
allows to share computations, which often makes a huge difference in practice.
In the same spirit, \sklearn sometimes provides the
\texttt{fit\_predict} method for combining the \texttt{fit} and \texttt{predict}
methods. This is especially useful for retrieving cluster predictions with
respect to the training data \texttt{X\_train} in clustering algorithms.

@@

As an example, the following snippet illustrates how to create a complex worflow
computing both linear PCA and kernel PCA components on \texttt{X\_train}
(through a \texttt{FeatureUnion}), selecting the 100 best of those features and
then feeding them to an $\ell_2$-regularized logistic regression model:
\begin{pythoncode}
    from sklearn.pipeline import Pipeline
    from sklearn.decomposition import PCA, KernelPCA
    from sklearn.feature_selection import SelectKBest

    fu = FeatureUnion([('pca', PCA()),
                       ('kpca', KernelPCA(kernel="rbf"))])

    Pipeline([("pca", fu,
              ("feat_sel", SelectKBest(k=100)),
              ("log_reg", LogisticRegression(penalty="l2"))
    ]).fit(X_train, y_train).predict(X_test)
\end{pythoncode}

@@

As an example, the snippet below illustrates how to optimize the
hyper-parameters of an SVM classifier. It maximizes the $F_1$ score of the
classifier through a 10-fold cross-validation scheme.
\begin{pythoncode}
    from sklearn.svm import SVC
    from sklearn.grid_search import GridSearchCV

    param_grid = [
      {"C": [1, 10, 100, 1000], "kernel": ["linear"]},
      {"C": [1, 10, 100, 1000], "gamma": [0.001, 0.0001], 
       "kernel": ["rbf"]},
    ]

    clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
    clf.fit(X_train, y_train)
    y_pred = clf.predict(X_test)
\end{pythoncode}
Note that, in the above example, two distinct hyper-parameter grids are
considered. This is because it only makes sense to search for \texttt{gamma}
when \texttt{kernel} is set to \texttt{"rbf"}.  Additionally, we see that
\texttt{GridSearchCV} has a \texttt{predict} method, like any other classifier.
This is because \texttt{GridSearchCV} and \texttt{RandomizedSearchCV} in fact
delegate the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
\texttt{score} methods to the best estimator, thus mimicking its behavior.
